---
title: "HW3"
author: "Victoria"
date: "2024-05-14"
output: html_document
---

## Libraries
```{r}
library(vegan)
library(readxl)
library(ggplot2)
library(impute)
library(factoextra)
library(psych)
library(ggforce)
library(rstatix)
```


```{r}
df <- read_excel("Sleepy lizard.xlsx")
df <- df[, c(4:7,9, 13:18)]
df$Treatment <- as.factor(df$Treatment)
df$Habitat <- as.factor(df$Habitat)
df$Connectivity <- as.factor(df$Connectivity)
blood <- df[, c(5:11)]
summary(blood)
```


```{r}
boxplot(blood)
```
scaling
```{r}
blood <- scale(log(blood + 1), scale = FALSE)
boxplot(blood)
```
Check the equality of variance in groups for perMANOVA.
```{r}
dist_blood <- vegdist(blood, method  = "euclidean")
PCO_blood <- betadisper(dist_blood, df$Treatment)
plot(PCO_blood)
anova(PCO_blood)
```
```{r}
dist_blood <- vegdist(blood, method  = "euclidean")
PCO_blood <- betadisper(dist_blood, df$Habitat)
plot(PCO_blood)
anova(PCO_blood)
```

```{r}
dist_blood <- vegdist(blood, method  = "euclidean")
PCO_blood <- betadisper(dist_blood, df$Connectivity)
plot(PCO_blood)
anova(PCO_blood)
```
no significant difference of dispersion between groups

Task 1

a) 
```{r}
adonis2(blood ~ df$Treatment, method = "euclidean")
```
significant

b)
```{r}
df_mod <- blood %>%
    as.data.frame() %>%
    filter(df$Treatment == 2)
habitat_mod <- df[df$Treatment == 2, ]$Habitat
adonis2(df_mod ~ habitat_mod, method = "euclidean")
```
not significant

c) 
```{r}
connectivity_mod <- df[df$Treatment == 2, ]$Connectivity
adonis2(df_mod ~ connectivity_mod, method = "euclidean")
```
not significant

Task 2
```{r}
adonis2(df_mod ~ connectivity_mod + habitat_mod, method = "euclidean")
```
blood composition ~ Habitat * Connectivity not significant

# Conclusion
perMANOVA analysis shows the significant difference in blood composition of lizards from unmodified and heavily modified landscapes. 
For degree of connectivity and habitat type there is no significant difference in blood for heavily modified landscapes.